yield response function

### 2025-09-30 codes/R/visualized_results.qmd
Codes modified (#fields added)

# ---- Constants: all 18 result folders ----
MODEL_DIRS <- c(
  "CF_outcome_1_fields","CF_outcome_5_fields","CF_outcome_10_fields",
  "RANN_outcome_one_field","RANN_outcome_five_fields","RANN_outcome_ten_fields",
  "Simple_ANN_outcome_one_field","Simple_ANN_outcome_two_fields","Simple_ANN_outcome_three_fields", "Simple_ANN_outcome_four_fields",
  "Simple_ANN_outcome_five_fields","Simple_ANN_outcome_ten_fields","Simple_ANN_outcome_twenty_fields",
  "NO_RF_one_field","NO_RF_five_fields","NO_RF_ten_fields",
  "HalfRRF_one_field","HalfRRF_five_fields","HalfRRF_ten_fields",
  "half_RANN_outcome_one_field","half_RANN_outcome_five_fields","half_RANN_outcome_ten_fields", "half_RANN_outcome_twenty_fields"
)

@@

# ---- function ----
rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))

parse_n_fields <- function(folder) {
  case_when(
    str_detect(folder, "(?:_1_fields|one_field)\\b")      ~ 1L,
    str_detect(folder, "(?:_2_fields|two_fields?)\\b")    ~ 2L,
    str_detect(folder, "(?:_3_fields|three_fields?)\\b")  ~ 3L,
    str_detect(folder, "(?:_4_fields|four_fields?)\\b")   ~ 4L,
    str_detect(folder, "(?:_5_fields|five_fields?)\\b")   ~ 5L,
    str_detect(folder, "(?:_10_fields|ten_fields?)\\b")   ~ 10L,
    str_detect(folder, "(?:_20_fields|twenty_fields?)\\b")~ 20L,
    TRUE ~ NA_integer_
  )
}

parse_model <- function(folder) {
  f <- str_to_lower(folder)
  case_when(
    str_starts(f, "cf_")             ~ "CF",
    str_starts(f, "rann_outcome")    ~ "RANN",
    str_starts(f, "half_rann")       ~ "HalfRANN",
    str_starts(f, "simple_ann")      ~ "Simple_ANN",
    str_starts(f, "no_rf")           ~ "NO_RF",
    str_starts(f, "halfrrf")         ~ "HalfRRF",
    TRUE                             ~ folder
  )
}

# Read any object: RDS, CSV
read_any <- function(f) {
  ext <- tolower(path_ext(f))
  if (ext == "csv") return(read_csv(f, show_col_types = FALSE))
  obj <- tryCatch(readRDS(f), error = function(e) NULL)
  if (!is.null(obj)) return(obj)
  env <- new.env(parent = emptyenv())
  ok <- tryCatch({ nm <- load(f, envir = env); length(nm) > 0 }, error = function(e) FALSE)
  if (ok) return(env[[ls(env)[1]]])
  stop("Unsupported or unreadable file: ", f)
}

# Standardize to two numeric vectors (per-sim rows)
read_eonr_file <- function(f) {
  obj <- read_any(f)
  if (is.list(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF
  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF list
  } else {
    stop("Unknown structure (need pred/true or opt_N_hat/opt_N): ", f)
  }
}

# 
files_in_dir <- function(dir_path) {
  dir_ls(dir_path, type = "file", recurse = TRUE) %>%
    # Names like sim_055_eonr or EONR_1 
    keep(~ str_detect(path_file(.x),
           regex("^(sim_\\d+(?:_eonr)?|EONR_\\d+)(?:\\.(rds|rda|rdata|csv))?$",
                 ignore_case = TRUE))) %>%
    # drop anything starting with "validation"
    discard(~ str_detect(path_file(.x), regex("^validation", ignore_case = TRUE))) %>%
    # drop macOS Icon files
    discard(~ grepl("Icon", basename(.x), fixed = TRUE))
}

# get sim id from filename base
extract_sim_id <- function(base) {
  m <- str_match(base, regex("^(?:sim_(\\d+)(?:_eonr)?|EONR_(\\d+))", ignore_case = TRUE))
  coalesce(as.integer(m[,2]), as.integer(m[,3]))
}

# 
row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
  tryCatch({
    base <- path_ext_remove(path_file(f))
    sim_id <- extract_sim_id(base)
    if (is.na(sim_id)) return(NULL)

@@

# ---- create table  ----
build_rmse_table <- function(base_dir = here("results"),
                             include_dirs = MODEL_DIRS,
                             expected_rows = EXPECTED_ROWS) {
  dirs <- file.path(base_dir, include_dirs)
  dirs <- dirs[dir_exists(dirs)]

@@

    map_dfr(files, row_from_file, model = model, n_f = n_f,
            expected_rows = expected_rows)
  }) %>%
    filter(!is.na(sim), !is.na(`#field`), !is.na(RMSE)) %>%
    arrange(Model, `#field`, sim)
}

@@

```{r}
# ---- run it (all models) ----
rmse_tbl_raw <- build_rmse_table(
  base_dir = here("results"),
  include_dirs = MODEL_DIRS,
  expected_rows = EXPECTED_ROWS
)

@@

#  model names: row labels in the graph
model_to_row <- c(
  "Simple_ANN" = "No_ANN",
  "RANN"       = "DO_ANN",
  "NO_RF"      = "NO_RF",
  "HalfRANN" = "SO_ANN",
  "HalfRRF"  = "SO_RF",
   "CF"       = "CF"
)

rmse_plot_df <- rmse_tbl_raw %>%
  filter(ok_rows, Model %in% names(model_to_row)) %>%
  mutate(
    row_lab = recode(Model, !!!model_to_row),
    col_lab = factor(`#field`,
                     levels = c(1, 2, 3, 4, 5, 10, 20),
                     labels = c("Number of fields: 1",
                                "Number of fields: 2",
                                "Number of fields: 3",
                                "Number of fields: 4",
                                "Number of fields: 5",
                                "Number of fields: 10",
                                "Number of fields: 20"))
  )

#  check 
rmse_plot_df %>% count(row_lab, col_lab)

# --- setup ---
xmin <- 0
nice_ceiling <- function(x, base = 5) base * ceiling(x / base)
xmax <- nice_ceiling(max(rmse_plot_df$RMSE, na.rm = TRUE), base = 5)

@@

# --- Plot ---
p <- ggplot(rmse_plot_df, aes(x = RMSE)) +
  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left") +
  facet_grid(row_lab ~ col_lab, scales = "fixed") +  
  scale_x_continuous(limits = c(xmin, xmax),
                     breaks = seq(xmin, xmax, length.out = 6),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "RMSE by iteration", y = "Count") +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 14)
  )

p

# ggsave(here::here("results", "rmse_hist_3rows.png"), p, width = 12, height = 7, dpi = 300)

@@

# 
rmse_df <- rmse_tbl_raw %>%
  dplyr::group_by(Model, `#field`, sim) %>%  
  dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
  # Rename models
  dplyr::mutate(
    Model = dplyr::recode(
      Model,
      "Simple_ANN" = "NO_ANN",
      "HalfRANN"   = "SO_ANN",
      "RANN"       = "DO_ANN",
      "HalfRRF"    = "SO_RF",
      "CF"         = "CF",
      "NO_RF"      = "NO_RF",
      .default     = Model
    ),
    field_f = factor(`#field`, levels = c(1, 2, 3, 4, 5, 10, 20),
                     labels = c("1 field", "2 fields", "3 fields", "4 fields",  "5 fields", "10 fields", "20 fields"))
  )

# 3) summary stats 
rmse_summary <- rmse_df %>%
  dplyr::group_by(Model, field_f) %>%
  dplyr::summarise(
    n      = dplyr::n(),
    mean   = mean(RMSE, na.rm = TRUE),
    median = median(RMSE, na.rm = TRUE),
    sd     = sd(RMSE, na.rm = TRUE),
    se     = sd / sqrt(n),
    lower  = mean + qt(0.025, df = n - 1) * se,
    upper  = mean + qt(0.975, df = n - 1) * se,
    .groups = "drop"
  )

@@

# Best model by order

@@

best_models %>%
  gt() %>%
  tab_header(
    title = "Best Models by Field Count",
    subtitle = "Ordered by lowest mean RMSE"
  ) %>%
  fmt_number(
    columns = c(mean),
    decimals = 3
  ) %>%
  cols_label(
    field_f = "# Fields",
    mean    = "mean RMSE"

>>>>

@@

# --------
rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))

# number of fields 
parse_n_fields <- function(folder) {
  f <- str_to_lower(folder)

  num <- str_match(f, "(_|\\b)(\\d+)_?fields?\\b")[,3]
  if (!is.na(num)) return(as.integer(num))

  words <- c("one"=1,"two"=2,"three"=3,"four"=4,"five"=5,
             "ten"=10,"twenty"=20,"thirty"=30,"forty"=40)
  word <- str_match(f, "(one|two|three|four|five|ten|twenty|thirty|forty)_?fields?")[,2]
  if (!is.na(word)) return(words[[word]])

  NA_integer_
}

# --- Model name mapping ---
parse_model <- function(folder) {
  f <- str_to_lower(folder)

  case_when(
    str_detect(f, "^cf")          ~ "CF",
    str_detect(f, "^rann")        ~ "DO_ANN",
    str_detect(f, "^half_rann")   ~ "SO_ANN",
    str_detect(f, "^simple_ann")  ~ "No_ANN",
    str_detect(f, "^no_rf")       ~ "NO_RF",
    str_detect(f, "^halfrrf")     ~ "SO_RF",
    TRUE                          ~ NA_character_
  )
}

# read RDS, CSV
read_any <- function(f) {
  ext <- tolower(path_ext(f))
  if (ext == "csv") return(read_csv(f, show_col_types = FALSE))
  obj <- tryCatch(readRDS(f), error = function(e) NULL)
  if (!is.null(obj)) return(obj)
  env <- new.env(parent = emptyenv())
  ok <- tryCatch({ nm <- load(f, envir = env); length(nm) > 0 }, error = function(e) FALSE)
  if (ok) return(env[[ls(env)[1]]])
  stop("Unsupported or unreadable file: ", f)
}

read_eonr_file <- function(f) {
  obj <- read_any(f)
  if (is.list(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF
  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF list
  } else {
    stop("Unknown structure (need pred/true or opt_N_hat/opt_N): ", f)
  }
}

# list files in dir 
files_in_dir <- function(dir_path) {
  dir_ls(dir_path, type = "file", recurse = TRUE) %>%
    keep(~ str_detect(path_file(.x),
                      regex("^(sim_\\d+(?:_eonr)?|EONR_\\d+)(?:\\.(rds|rda|rdata|csv))?$",
                            ignore_case = TRUE))) %>%
    discard(~ str_detect(path_file(.x), regex("^validation", ignore_case = TRUE))) %>%
    discard(~ grepl("Icon", basename(.x), fixed = TRUE))
}

# extract sim id from filename
extract_sim_id <- function(base) {
  m <- str_match(base, regex("^(?:sim_(\\d+)(?:_eonr)?|EONR_(\\d+))", ignore_case = TRUE))
  coalesce(as.integer(m[,2]), as.integer(m[,3]))
}

row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
  tryCatch({
    base <- path_ext_remove(path_file(f))
    sim_id <- extract_sim_id(base)
    if (is.na(sim_id)) return(NULL)

@@

# --- RMSE table ----
build_rmse_table <- function(base_dir = here("results"),
                             expected_rows = EXPECTED_ROWS) {

  dirs <- fs::dir_ls(base_dir, type = "directory", recurse = FALSE) %>%
    keep(~ str_detect(basename(.x),
                      regex("^(cf|rann|half_rann|simple_ann|no_rf|halfrrf)",
                            ignore_case = TRUE)))

@@

    map_dfr(files, row_from_file, model = model, n_f = n_f,
            expected_rows = expected_rows)
  }) %>%
    filter(!is.na(Model), !is.na(`#field`), !is.na(RMSE)) %>%
    arrange(Model, `#field`, sim)
}

@@

```{r}
# ---- run it (all models) ----
rmse_tbl_raw <- build_rmse_table(
  base_dir = here("results"),
  expected_rows = EXPECTED_ROWS
)

@@

# -- labels for models ---
model_names <- function(models) {
  map <- c(
    "Simple_ANN" = "No_ANN",
    "RANN"       = "DO_ANN",
    "NO_RF"      = "NO_RF",
    "HALFRANN"   = "SO_ANN",
    "HALFRRF"    = "SO_RF",
    "CF"         = "CF"
  )
  # use mapping if available, else leave unchanged
  out <- unname(map[models])
  ifelse(is.na(out), models, out)
}

# --- plotting data ---
rmse_plot_df <- rmse_tbl_raw %>%
  filter(ok_rows) %>%
  mutate(
    row_lab = model_names(Model),
    col_lab = factor(
      `#field`,
      levels = sort(unique(`#field`)),   # discover field counts
      labels = paste0("Number of fields: ", sort(unique(`#field`)))
    )
  )

# --- Check ---
rmse_plot_df %>% count(row_lab, col_lab)

# ------
xmin <- 0
nice_ceiling <- function(x, base = 5) base * ceiling(x / base)
xmax <- nice_ceiling(max(rmse_plot_df$RMSE, na.rm = TRUE), base = 5)

@@

# --- Plot ---
p <- ggplot(rmse_plot_df, aes(x = RMSE)) +
  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left") +
  facet_grid(row_lab ~ col_lab, scales = "fixed") +  
  scale_x_continuous(
    limits = c(xmin, xmax),
    breaks = seq(xmin, xmax, length.out = 6),
    expand = c(0, 0)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "RMSE by iteration", y = "Count") +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 14)
  )

p
# ggsave(here::here("results", "rmse_hist_3rows.png"), p, width = 12, height = 7, dpi = 300)

@@

# --- Prepare rmse_df ---
rmse_df <- rmse_tbl_raw %>%
  dplyr::group_by(Model, `#field`, sim) %>%  
  dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
  dplyr::mutate(
    # Field labels (auto from unique values)
    field_f = factor(
      `#field`,
      levels = sort(unique(`#field`)),
      labels = paste0(sort(unique(`#field`)),
                      ifelse(sort(unique(`#field`)) == 1, " field", " fields"))
    )
  )

# --- Summary stats ---
rmse_summary <- rmse_df %>%
  dplyr::group_by(Model, field_f) %>%
  dplyr::summarise(
    n      = dplyr::n(),
    mean   = mean(RMSE, na.rm = TRUE),
    median = median(RMSE, na.rm = TRUE),
    sd     = sd(RMSE, na.rm = TRUE),
    se     = sd / sqrt(n),
    lower  = mean + qt(0.025, df = n - 1) * se,
    upper  = mean + qt(0.975, df = n - 1) * se,
    .groups = "drop"
  ) %>%
  # remove 2-field and 4-field combinations
  dplyr::filter(!field_f %in% c("2 fields", "4 fields"))

@@

# Best model by the numbers of fields

@@

best_models %>%
  gt() %>%
  tab_header(
    title = "best models by #field",
    subtitle = "Ordered by lowest mean RMSE"
  ) %>%
  fmt_number(
    columns = c(mean),
    decimals = 3
  ) %>%
  cols_label(
    field_f = "# fields",
    mean    = "mean RMSE"
